---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
#1. Load the data

data <- read.csv("activity.csv")

#2. Convert the class of the date column from factor to Date.

data$date <- as.Date(data$date)
```

## What is mean total number of steps taken per day?

1. Total number of steps taken per day

```{r}
ts <- aggregate(steps~date,data,sum)
ts
```

2. Histogram of the total number of steps taken each day

```{r}
par(mar=c(5, 4, 4, 2) + 0.1)
hist(ts$steps,main="Total Number of Steps Taken Each Day",xlab="Number of Steps")
```

3. The mean and median of the total number of steps taken per day

```{r}
mean(ts$steps)

median(ts$steps)
```



## What is the average daily activity pattern?

```{r}
as <- aggregate(steps~interval,data,mean)
plot(as$interval,as$steps,type="l",main="Average Number of Steps Taken for Each Interval",xlab="5-minute Interval",ylab="Average Number of Steps Taken")

maxsteps <- as$interval[which(as$steps==max(as$steps))]
```
The No. `r maxsteps` interval contains the maxium number of steps.

## Imputing missing values

The total number of missing values in the dataset is
```{r}
sum(is.na(data$steps))
```

Use the mean for the 5-minute interval to fill in all of the missing values

```{r}
newdata <- data
for (i in 1:nrow(newdata)) if(is.na(newdata$steps[i]))newdata$steps[i] <- as$steps[which(newdata$interval[i]==as$interval)]
```

Make a histogram of the total number of steps taken each day
```{r}
ts2 <- aggregate(steps~date,newdata,sum)
hist(ts2$steps,main="Total Number of Steps Taken Each Day After filling in all NAs",xlab="Number of Steps")
```

The mean and median total number of steps taken per day
```{r}
# The mean
mean(ts2$steps)

# The median
median(ts2$steps)
```

The mean doesn't change becuase I filled in means for each 5-minute interval. However, the median changed a bit, from 10765 to 10766.19.

## Are there differences in activity patterns between weekdays and weekends?
```{r,echo=TRUE}
# Weekday or weekend
library(chron)
library(plyr)
newdata <- mutate(newdata,day=is.weekend(date))
for (i in 1:nrow(newdata)) if(newdata$day[i]) newdata$day[i] <- "weekend" else newdata$day[i] <- "weekday"
invisible(as.factor(newdata$day))

# Make a panel plot
library(lattice)
mstep <- aggregate(newdata$steps,by=list(newdata$interval,newdata$day),mean)
names(mstep) <- c("interval","day","steps")
xyplot(steps~interval | day, mstep,type="l",layout=c(1,2),xlab="Interval",ylab="Number of steps")
```

Activity patterns differ between weekdays and weekends. During weekends, people walk more in the whole day. However, during weekdays, people walk more in the morning.